---
title: "Model Extension Report"
author: "Leonardo Feitosa"
date: "03/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(patchwork)
library(kableExtra)
```

# Introduction

  Invasive species management requires thorough understanding of life history characteristics, which can be used to model population dynamics and analyze effectiveness of various intervention strategies. Lionfish (*Pterois volitans/miles*) are an invasive species in the Caribbean, and have caused substantial harm to both the regional biological community and the people who rely on coral reefs for economic subsistence (Albins, 2015; Albins & Hixon, 2013). Since individual growth rates are often dependent on local environmental conditions, spatially-explicit estimates are necessary to better understand and manage lionfish invasions. Growth curves for this species have been generated for many locations throughout the Caribbean Sea (Akins et al., 2014; Rodríguez-Cortés, 2015). The paper we analyze in this project, by Eric G. Johnson and Mary Katherine Swenarton (2016), fills the knowledge gap on the individual-level growth patterns in a temperate population in Northeastern Florida. 

  Growth studies are also critical to understand how predator-prey interactions might be influenced by a rapid-growth invasive species such as the lionfish (Lorenzen, 2006). Johnson and Swenarton (2016) study the growth patterns and evaluate the population structure of *Pterois volitans/miles* through length-frequency data, applying a modified version of the von Bertalanffy Growth Function (VBGF) to include seasonality as an environmental parameter. In this paper, we recreate some of the analyses they performed in their paper and discuss the ecological importance of their results.  

# Model Presentation

The von Bertalanffy growth function uses length as a proxy for age. Generally, the length of a fish increases as it gets older, and many fish species demonstrate rapid growth as juveniles that then decreases as they near their maximum size. For species which experience seasonal temperature variations, such as temperate lionfish, the seasonalized version of this length-age function is often most appropriate:

$$
\begin{align}
\text{Seasonalized:} 
\newline 
L_t & = L_\infty [1-e^{-(K(t-t_0))-S(t)+S(t_0)}] \\ 
\newline
S(t)  &= \frac{CK}{2\pi} \cdot sin \pi(t-t_s)
\newline
S(t_0) &= \frac{CK}{2\pi} \cdot sin \pi(t_0-t_s)
\end{align}
$$

Here, the length of a fish at a certain age ($L_t$) is a function of the Brody growth coefficient ($K$) and fluctuations in the growth rate depending on the seasonal growth patterns (described by $S(t)$ and $S(t_0)$). The intensity of these fluctuations is controlled by $C$, where $C = 0$ indicates no seasonality, and $C = 1$ indicates complete seasonality. A complete list of parameters and their descriptions are included in Table 1. 


## Model Extension's Key Goals & Findings:

The paper analyzed in this project (Johnson & Swenarton 2016) consists of the calculations of growth for the lionfish (Pterois volitans/miles) in Northeastern Florida. While growth curves for this species have been generated throughout the region independently (Akins et al., 2014; Rodríguez-Cortés, 2015), the paper we analyzed does it in a very local scale within a limited time frame (2013-2015). Therefore, my proposal for an extension of this model is to include sea water temperature as another parameter, given that water temperature is key to the development and reproductive parameters of fish (Rountrey et al., 2014). Furthermore, given the recent extension records of the invasive lionfish beyond the Caribbean and as far south as Rio de Janeiro, southeastern Brazil (Ferreira et al., 2015; Luiz et al., 2021), the populations of this species are likely to be subjected to different sea water temperatures, thus likely exhibiting different growth patterns.

## Methods

#### Sampling

Sampling occurred from 2013-2015 during "lionfish roundup" tournaments and other opportunistic sampling via spearfishing. In the lab, lionfish were measured for standard and total length, weighed, and sexed. The authors retrieved otoliths from 100 lionfish to determine age directly. 

#### Model Creation

- Length-frequency histograms were used for lengths 0-450mm in 10mm increments.
- Only used data from 2014 due to temporal resolution

About the model:

The von Bertalanffy growth function uses length as a proxy for age. For species which experience seasonal temperature variations, such as temperate lionfish, the seasonalized version of this function is often most appropriate:

$$
\begin{align}
\text{Seasonalized:} 
\newline 
L_t & = L_{inf} [1-e^{-(K(t-t_0))-S(t)+S(t_0)}] \\ 
\newline
S(t)  &= \frac{CK}{2\pi} \cdot sin \pi(t-t_s)
\newline
S(t_0) &= \frac{CK}{2\pi} \cdot sin \pi(t_0-t_s)
\end{align}
$$

$L_t$ = length of fish at age $t$
$L_inf$ = asymptotic maximum length
$K$ = Brody growth coefficient
$t_0$ = theoretical time at which fish was length zero
$t_s$ = seasonal growth oscillations relative to $t_0$
$t_w = t_s + 0.5$ = winter point, time of slowest growth
$C$ = intensity of seasonal growth oscillation, ($0 \leq C \leq 1$ where zero = no seasonality)

Included measure of variance of size-at-age as a parameter, and estimated the proportion of lionfish in each age class during each sampling month and year:

$$
\begin{align}
n_{a,i,t} = N_t \cdot P_{a,t} \cdot P(L_{lower} \leq L_i \leq L_{upper} | N(\bar{L_{a,t}},\sigma_{a}^{2}))
\end{align}
$$
$N_t$ = total number of individuals captured in month $t$
$P_{a,t}$ = probability of fish captured in month $t$ being age $a$
$L_{lower}$ and $L_{upper}$ = lower and upper bounds of predcted 10mm size bin
$N(\bar{L_{a,t}}, sigma_a)$ = normal probability density function with mean length $\bar{L_{a,t}}$ of fish of age $a$ in month $t$ as estimated from seasonalized VBGF and model esimated standard deviation at age, $sigma_a$. 

Size distributions overlap across ages, so total expected fish of size $i$ in month $t$ calculated by summing expected contributions to size class $i$ from each age:
$N_{i,t} = \sum_{a=0}^{3}n_{i,a,t}$

Uses maximum likelihood approach to fit a predicted distribution to the length-frequency histogram - length is a proxy for age, so the model estimates the proportion of fish of a given size in each age class. 

Best fit model:
$$
\begin{align}
L_t = 448 [1-e^{-0.47t+[\frac{0.061*0.47}{2\pi}]sin\pi(t-0.71)]-[\frac{0.61*0.47}{2\pi}sin\pi(0.71)]}]
\end{align}
$$

***Table 1. Description of variables and parameters for the seasonalized von-Bertalanffy growth equation***
```{r echo=FALSE}
# Insert table here
table <- data.frame("Term" = c("$L_t$",
                               "$t$",
                               "$L_\\infty$",
                               "$K$",
                               "$t_0$",
                               "$t_s$",
                               "$t_w$",
                               "$t_b$",
                               "$C$",
                               "$T$"),
                    "Estimate" = c("output variable (mm)",
                                   "input variable (yr)",
                                   "448 mm",
                                   "0.47 yr$^{-1}$",
                                   "0 yr",
                                   "0.71 yr",
                                   "not stated",
                                   "0.41 yr",
                                   "0.61",
                                   "22.5(Fl), 23 (BRJ), 24 (BFN), 25 (BNB)"),
                    "Description" = c("length of fish at age $t$",
                                      "age $t$",
                                      "asymptotic maximum length",
                                      "Brody growth coefficient",
                                      "theoretical time at which fish was length zero artificial constant set at zero for this paper)",
                                      "seasonal growth oscillation relative to $t_0$",
                                      "winter point, time of slowest growth ($t_w = t_s + 0.5$)",
                                      "time of annual recruitment",
                                      "intensity of seasonal growth oscillation 
                                      ($0 \\leq C \\leq 1$, where zero indicates 
                                      no seasonality and one indicates complete seasonality)",
                                      "water temperature (ºC). FL - Florda, BRJ - Rio de Janeiro,
                                      BFN - Fernando de Noronha Archipelago, BNB - Northern Brazilian coast"))
kable(table) %>% 
  kable_styling()
```

## Replication

### Data Processing

```{r}
lionfish_tl <- read.csv(here("data", "lionfish_length-frequency_data.csv"), sep = ";") %>% 
  clean_names() %>% 
  select(total_length_tl:august_2015)

# Create tibble in long format and separate the original year column into month and year
lionfish_long <- lionfish_tl %>% 
  pivot_longer(cols = april_2013:august_2015,
               names_to = "year",
               values_to = "frequency") %>% 
  separate(year, c("month", "year"), "_")

# Sum the frequencies per year
lionfish_sum <- lionfish_long %>% 
  group_by(year, total_length_tl) %>% 
  summarise(freq_sum = sum(frequency))

```

##### Cleaner Alternative: ggplot + facet_wrap()

```{r}
ggplot(data = lionfish_long) +
  geom_col(aes(x = total_length_tl, y = frequency)) +
  theme_classic() +
  labs(x = "Total Length",
       y = "Frequency") +
  facet_wrap(~year, scales = "free")
```

```{r parameters}
l_inf = 448 # asymptotic maximum length
k = 0.47    # Brody growth coefficient
c = 0.61    # intensity of seasonal growth oscillation. 
               # varies from 0 to 1, with zero being no seasonality 
               # and one extreme seasonality
t_0 = 0    # theoretical time when fish was length zero
t_s = 0.71 # timing of seasonal growth oscillation...
              # slowest growth is at winter point, tw = ts + 0.5

# Values of temperature for each location
# Maranhão
t_m = 25  # Moura et al (2016)

# Fernando de Noronha
t_fn = 24 # Prazeres et al (2012)

# Rio de Janeiro
t_rj = 23 # Lima & Coutinho (2016)

# Northeast Florida
t_fl = 22.5 # based on monthly averages for Jacksonville obtained from https://seatemperature.info/jacksonville-water-temperature.html 

```

## Replicating the best fit model used in Johnson & Swenarton (2016)

### Figure 4

```{r}
# Parameters 
l_inf = 448 # asymptotic maximum length
k = 0.47    # Brody growth coefficient
c = 0.61    # intensity of seasonal growth oscillation. 
               # varies from 0 to 1, with zero being no seasonality 
               # and one extreme seasonality
t_0 = 0    # theoretical time when fish was length zero
t_s = 0.71 # timing of seasonal growth oscillation...
              # slowest growth is at winter point, tw = ts + 0.5

# Set up loop to iterate model across time/age
tset <- seq(from = 0, to = 4, by = 0.001)
lf <- data.frame(tset, lengths = NaN*tset)

for(i in 1:length(tset)){
  t <- tset[i]
  l_t <- l_inf*(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
  lf$lengths[i] <- l_t
}

# Plot the model output
ggplot() +
  geom_line(data = lf,
            aes(x = tset, y = lengths)) +
  labs(x = "Age (Years)",
       y = "Total Length (mm)") +
  theme_classic()
```

##  Figure 3

Now need to try to use the growth to estimate the age structure of the population:


```{r}
# Starting age structure,
lf_ap13 <- lionfish_long %>% 
  filter(month == "april" & year == "2013") %>% 
  uncount(frequency) %>% 
  select(total_length_tl) %>% 
  rename(starting_length = total_length_tl) 

# Create table of age-length relationships
# (Nice alternative to solving the model for length)
lf_nice <- lf %>%  
  mutate(length_round = round(lengths, digits = 0))%>% 
  group_by(length_round) %>% 
  slice(1) %>% 
  ungroup()

lf_sim <- data.frame(start_l = lf_ap13$starting_length,
                     start_a    = NaN*length(lf_ap13),
                     a_1 = NaN*length(lf_ap13),
                     l_1 = NaN*length(lf_ap13),
                     a_2 = NaN*length(lf_ap13),
                     l_2 = NaN*length(lf_ap13), 
                     a_3 = NaN*length(lf_ap13),
                     l_3 = NaN*length(lf_ap13),
                     a_4 = NaN*length(lf_ap13),
                     l_4 = NaN*length(lf_ap13))

# Test out first as a loop
for(i in 1:length(lf_sim$start_l)){
    # Starting Age/Length
    a <- lf_nice$tset[lf_nice$length_round == lf_sim$start_l[i]] 
    lf_sim$start_a[i] <- a
    
    # 1-yr Age/Length
    t <- a + 1
    lf_sim$a_1[i] <- t
    lf_sim$l_1[i] <- l_inf*(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
    
    # 2-yr Age/Length
    t <- t + 1
    lf_sim$a_2[i] <- t
    lf_sim$l_2[i] <- l_inf*(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
   
    # 3-yr Age/Length
    t <- t + 1
    lf_sim$a_3[i] <- t
    lf_sim$l_3[i] <- l_inf*(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
    
    # 3-yr Age/Length
    t <- t + 1
    lf_sim$a_4[i] <- t
    lf_sim$l_4[i] <- l_inf*(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
}


  
# Works!
ggplot(data = lf_sim) +
  geom_histogram(aes(x = start_l, fill = "April 2013"), alpha = 0.5)+
  geom_histogram(aes(x = l_1, fill = "April 2013 aged to April 2014"), alpha = 0.5) +
  geom_histogram(aes(x = l_2, fill = "April 2013 aged to April 2015"), alpha = 0.5) +
  geom_histogram(aes(x = l_3, fill = "April 2013 aged to April 2016"), alpha = 0.5) +
  geom_histogram(aes(x = l_4, fill = "April 2013 aged to April 2017"), alpha = 0.5) +
  labs(x = "length", y = "frequency", fill = NULL) +
  theme_classic()
```


## Model extension 

### Figure 3

```{r}
# Starting age structure,
lf_ap13 <- lionfish_long %>% 
  filter(month == "april" & year == "2013") %>% 
  uncount(frequency) %>% 
  select(total_length_tl) %>% 
  rename(starting_length = total_length_tl) 

# Create table of age-length relationships
# (Nice alternative to solving the model for length)
lf_nice <- lf %>%  
  mutate(length_round = round(lengths, digits = 0))%>% 
  group_by(length_round) %>% 
  slice(1) %>% 
  ungroup()
```

```{r for loop for florida}
lf_sim_fl <- data.frame(start_l_fl = lf_ap13$starting_length,
                     start_a_fl  = NaN*length(lf_ap13),
                     a_1_fl = NaN*length(lf_ap13),
                     l_1_fl = NaN*length(lf_ap13),
                     a_2_fl = NaN*length(lf_ap13),
                     l_2_fl = NaN*length(lf_ap13), 
                     a_3_fl = NaN*length(lf_ap13),
                     l_3_fl = NaN*length(lf_ap13),
                     a_4_fl = NaN*length(lf_ap13),
                     l_4_fl = NaN*length(lf_ap13))

# Test out first as a loop
for(i in 1:length(lf_sim_fl$start_l_fl)){
    # Starting Age/Length
    a_fl <- lf_nice$tset[lf_nice$length_round == lf_sim_fl$start_l_fl[i]] 
    lf_sim_fl$start_a[i] <- a_fl
    
    # 1-yr Age/Length
    t_flo <- a_fl + 1
    lf_sim_fl$a_1_fl[i] <- t_flo
    lf_sim_fl$l_1_fl[i] <- l_inf*t_fl*(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
    
    # 2-yr Age/Length
    t_flo <- t_flo + 1
    lf_sim_fl$a_2_fl[i] <- t_flo
    lf_sim_fl$l_2_fl[i] <- l_inf*t_fl*(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
   
    # 3-yr Age/Length
    t_flo <- t_flo + 1
    lf_sim_fl$a_3_fl[i] <- t_flo
    lf_sim_fl$l_3_fl[i] <- l_inf*t_fl*(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
    
    # 3-yr Age/Length
    t_flo <- t_flo + 1
    lf_sim_fl$a_4_fl[i] <- t_flo
    lf_sim_fl$l_4_fl[i] <- l_inf* t_fl*(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
}

```


```{r for loop for maranhao}
lf_sim_m <- data.frame(start_l_m = lf_ap13$starting_length,
                     start_a_m = NaN*length(lf_ap13),
                     a_1_m = NaN*length(lf_ap13),
                     l_1_m = NaN*length(lf_ap13),
                     a_2_m = NaN*length(lf_ap13),
                     l_2_m = NaN*length(lf_ap13), 
                     a_3_m = NaN*length(lf_ap13),
                     l_3_m = NaN*length(lf_ap13),
                     a_4_m = NaN*length(lf_ap13),
                     l_4_m = NaN*length(lf_ap13),
                     region = NaN*length(lf_ap13))

for(i in 1:length(lf_sim_m$start_l_m)){
    # Starting Age/Length
    a_m <- lf_nice$tset[lf_nice$length_round == lf_sim_m$start_l_m[i]] 
    lf_sim_m$start_a_m[i] <- a_m
    
    # 1-yr Age/Length
    t <- a_m + 1
    lf_sim_m$a_1_m[i] <- t
    lf_sim_m$l_1_m[i] <- l_inf* t_m *(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
    
    # 2-yr Age/Length
    t <- t + 1
    lf_sim_m$a_2_m[i] <- t
    lf_sim_m$l_2_m[i] <- l_inf*t_m *(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
   
    # 3-yr Age/Length
    t <- t + 1
    lf_sim_m$a_3_m[i] <- t
    lf_sim_m$l_3_m[i] <- l_inf*t_m *(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
    
    # 3-yr Age/Length
    t <- t + 1
    lf_sim_m$a_4_m[i] <- t
    lf_sim_m$l_4_m[i] <- l_inf*t_m *(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
    
    # add region to the df
    region <- t + 1
    lf_sim_m$region[i] <- c("MA")
}

```

```{r for loop for noronha}
lf_sim_fn <- data.frame(start_l_fn = lf_ap13$starting_length,
                     start_a_fn = NaN*length(lf_ap13),
                     a_1_fn = NaN*length(lf_ap13),
                     l_1_fn = NaN*length(lf_ap13),
                     a_2_fn = NaN*length(lf_ap13),
                     l_2_fn = NaN*length(lf_ap13), 
                     a_3_fn = NaN*length(lf_ap13),
                     l_3_fn = NaN*length(lf_ap13),
                     a_4_fn = NaN*length(lf_ap13),
                     l_4_fn = NaN*length(lf_ap13),
                     region = NaN*length(lf_ap13))

for(i in 1:length(lf_sim_fn$start_l_fn)){
    # Starting Age/Length
    a_fn <- lf_nice$tset[lf_nice$length_round == lf_sim_fn$start_l_fn[i]] 
    lf_sim_fn$start_a_fn[i] <- a_fn
    
    # 1-yr Age/Length
    t_f <- a_fn + 1
    lf_sim_fn$a_1_fn[i] <- t_f
    lf_sim_fn$l_1_fn[i] <- l_inf* t_fn *(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
    
    # 2-yr Age/Length
    t_f <- t_f + 1
    lf_sim_fn$a_2_fn[i] <- t_f
    lf_sim_fn$l_2_fn[i] <- l_inf*t_fn *(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
   
    # 3-yr Age/Length
    t_f <- t_f + 1
    lf_sim_fn$a_3_fn[i] <- t_f
    lf_sim_fn$l_3_fn[i] <- l_inf*t_fn *(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
    
    # 3-yr Age/Length
    t_f <- t_f + 1
    lf_sim_fn$a_4_fn[i] <- t_f
    lf_sim_fn$l_4_fn[i] <- l_inf*t_fn *(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
    
    # include region in the df
    region <- t_f + 1
    lf_sim_fn$region[i] <- c("FN")
}
```

```{r for loop for rj}
lf_sim_rj <- data.frame(start_l_rj = lf_ap13$starting_length,
                     start_a_rj = NaN*length(lf_ap13),
                     a_1_rj = NaN*length(lf_ap13),
                     l_1_rj = NaN*length(lf_ap13),
                     a_2_rj = NaN*length(lf_ap13),
                     l_2_rj = NaN*length(lf_ap13), 
                     a_3_rj = NaN*length(lf_ap13),
                     l_3_rj = NaN*length(lf_ap13),
                     a_4_rj = NaN*length(lf_ap13),
                     l_4_rj = NaN*length(lf_ap13),
                     region = NaN*length(lf_ap13))

for(i in 1:length(lf_sim_rj$start_l_rj)){
    # Starting Age/Length
    a_rj <- lf_nice$tset[lf_nice$length_round == lf_sim_rj$start_l_rj[i]] 
    lf_sim_rj$start_a_rj[i] <- a_rj
    
    # 1-yr Age/Length
    t_r <- a_rj + 1
    lf_sim_rj$a_1_rj[i] <- t_r
    lf_sim_rj$l_1_rj[i] <- l_inf* t_rj *(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
    
    # 2-yr Age/Length
    t_r <- t_r + 1
    lf_sim_rj$a_2_rj[i] <- t_r
    lf_sim_rj$l_2_rj[i] <- l_inf*t_rj *(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
   
    # 3-yr Age/Length
    t_r <- t_r + 1
    lf_sim_rj$a_3_rj[i] <- t_r
    lf_sim_rj$l_3_rj[i] <- l_inf*t_rj *(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
    
    # 3-yr Age/Length
    t_r <- t_r + 1
    lf_sim_rj$a_4_rj[i] <- t_r
    lf_sim_rj$l_4_rj[i] <- l_inf*t_rj *(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
    
    # include region in the df
    region <- t_r + 1
    lf_sim_rj$region[i] <- c("RJ")
}
```


```{r plot for figure 3 model extension}
ggplot(data = lf_sim_m) +
  geom_histogram(aes(x = start_l_m, fill = "April 2013"), alpha = 0.5)+
  geom_histogram(aes(x = l_1_m, fill = "April 2013 aged to April 2014"), alpha = 0.5) +
  geom_histogram(aes(x = l_2_m, fill = "April 2013 aged to April 2015"), alpha = 0.5) +
  geom_histogram(aes(x = l_3_m, fill = "April 2013 aged to April 2016"), alpha = 0.5) +
  geom_histogram(aes(x = l_4_m, fill = "April 2013 aged to April 2017"), alpha = 0.5) +
  labs(x = "Length", y = "Frequency", fill = NULL, title = "Simulation for Maranhao") +
  theme_classic()

ggplot(data = lf_sim_fn) +
  geom_histogram(aes(x = start_l_fn, fill = "April 2013"), alpha = 0.5)+
  geom_histogram(aes(x = l_1_fn, fill = "April 2013 aged to April 2014"), alpha = 0.5) +
  geom_histogram(aes(x = l_2_fn, fill = "April 2013 aged to April 2015"), alpha = 0.5) +
  geom_histogram(aes(x = l_3_fn, fill = "April 2013 aged to April 2016"), alpha = 0.5) +
  geom_histogram(aes(x = l_4_fn, fill = "April 2013 aged to April 2017"), alpha = 0.5) +
  labs(x = "Length", y = "Frequency", fill = NULL, title = "Simulation for Fernando de Noronha") +
  theme_classic()

ggplot(data = lf_sim_rj) +
  geom_histogram(aes(x = start_l_rj, fill = "April 2013"), alpha = 0.5)+
  geom_histogram(aes(x = l_1_rj, fill = "April 2013 aged to April 2014"), alpha = 0.5) +
  geom_histogram(aes(x = l_2_rj, fill = "April 2013 aged to April 2015"), alpha = 0.5) +
  geom_histogram(aes(x = l_3_rj, fill = "April 2013 aged to April 2016"), alpha = 0.5) +
  geom_histogram(aes(x = l_4_rj, fill = "April 2013 aged to April 2017"), alpha = 0.5) +
  labs(x = "Length", y = "Frequency", fill = NULL, title = "Simulation for Rio de Janeiro") +
  theme_classic()

ggplot(data = lf_sim_fl) +
  geom_histogram(aes(x = start_l_fl, fill = "April 2013"), alpha = 0.5)+
  geom_histogram(aes(x = l_1_fl, fill = "April 2013 aged to April 2014"), alpha = 0.5) +
  geom_histogram(aes(x = l_2_fl, fill = "April 2013 aged to April 2015"), alpha = 0.5) +
  geom_histogram(aes(x = l_3_fl, fill = "April 2013 aged to April 2016"), alpha = 0.5) +
  geom_histogram(aes(x = l_4_fl, fill = "April 2013 aged to April 2017"), alpha = 0.5) +
  labs(x = "length", y = "frequency", fill = NULL, title = "Florida Simulation") +
  theme_classic()
```


### Figure 4 - Model extension Leo

```{r}
# Set up loop to iterate model across time/age
tset <- seq(from = 0, to = 4, by = 0.001)
lf <- data.frame(tset, lengths = NaN*tset)
lf_m <- data.frame(tset, lengths = NaN*tset)
lf_fn <- data.frame(tset, lengths = NaN*tset)
lf_rj <- data.frame(tset, lengths = NaN*tset)
for(i in 1:length(tset)){
  t <- tset[i]
  
  l_t <- l_inf*t_fl *(1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
  l_t_m <- l_inf*t_m* (1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
  l_t_fn <- l_inf*t_fn* (1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
  l_t_rj <- l_inf*t_rj* (1-exp(-k*(t-t_0)+(((c*k)/(2*pi))*sinpi(t-t_s))-(((c*k)/(2*pi))*sinpi(t_s))))
  
  lf$lengths[i] <- l_t
  lf_m$lengths[i] <- l_t_m
  lf_fn$lengths[i] <- l_t_fn
  lf_rj$lengths[i] <- l_t_rj
}

# Plot the model output
ggplot() +
  geom_line(data = lf,
            aes(x = tset, y = lengths)) +
  geom_line(data = lf_m,
            aes(x = tset, y = lengths),
            color = "coral2") +
  geom_line(data = lf_fn,
            aes(x = tset, y = lengths),
            color = "blue4") +
  geom_line(data = lf_rj,
            aes(x = tset, y = lengths),
            color = "goldenrod2") +
  labs(x = "Age (Years)",
       y = "Total Length (mm)") +
  theme_classic()
```


